A Novel Sampling Method Using Stochastic Gradient Descent

### 2016-12-04 paper/main.tex

\newtheorem{conjecture}{Conjecture}

@@

\begin{conjecture}
For any $p(\bx)\in \D(\R^p)$, it is the stationary distribution of iterates in Algorithm 1 with parameter $(-\log p(\bx), \epsilon, \epsilon)$. 
\end{conjecture}
\begin{proof}
Let's verify this theorem by checking the detailed balance condition. Denote $g(\by|\bx)$ to be the density function of the transition, because $\by = \bx - \alpha \nabla f(\bx^k) + \epsilon^t$ and $\epsilon^t\sim \N(0, \sigma^2)$, we know
\begin{equation}
g(\by|\bx) = (2\pi \sigma^2)^{-p/2} \exp(- \frac{\|\by - \bx + \alpha \nabla h(\bx)\|^2}{2\sigma^2}).
\end{equation}
Then the detailed balance condition is 
\begin{equation}\label{Eq:detailBalance}
p(\bx^{k+1}) g(\bx^{k}|\bx^{k+1}) = p(\bx^{k}) g(\bx^{k+1}|\bx^{k}).
\end{equation}
Since we will only consider $\bx^k$ and $\bx^{k+1}$ in the proof. For notation ease, we denote $\bx^{k+1}$ as $\bx^+$. 
In order to prove \ref{Eq:detailBalance}, we would like to show the ratio between LHS and RHS to be 1. To be more specific, we have (Recall that $\sigma^2 = \alpha$)
\begin{eqnarray}
\frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = & \exp(-h(\bx^{+}) - \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} + h(\bx^{k}) + \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha}).
\end{eqnarray}
Note that 
\[h(\bx^+) - h(\bx^k) = \frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + O(\|\bx^+ - \bx^k\|^2),
\]
and 
\begin{eqnarray}
 & \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} - \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha} \\
 =& -\frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + \alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2.
\end{eqnarray}
Here, $\|\bx^+ - \bx^k\|$ will be roughly no bigger than $\alpha$ because $\bx^+ = \bx^k - \alpha \nabla h(\bx^k) + \epsilon^k$.  Thus it means $O(\|\bx^+ - \bx^k\|) = O(\alpha^2)$. What's more, $\alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2 = O(\alpha^2)$. 
Thus we know 
\begin{eqnarray}
\log \frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = &O(\alpha^2).
\end{eqnarray}
Therefore, we know the detailed balance equality holds approximately especially when $\alpha$ is small.

When $h$ is a quadratic function, this equation holds exactly. 
\end{proof}

@@

Some figures...

>>>>

\newtheorem{conjecture}{Conjecture}
\newtheorem{theorem}{Theorem}
\newtheorem{experiment}{Experiment}

@@

\begin{conjecture}
For any $p(\bx)\in \D(\R^p)$, it is the stationary distribution of iterates in Algorithm 1 with parameter $(-\log p(\bx), \epsilon, 2\epsilon)$. 
\end{conjecture}
\begin{proof}
Let's verify this theorem by checking the detailed balance condition. Denote $g(\by|\bx)$ to be the density function of the transition, because $\by = \bx - \alpha \nabla f(\bx^k) + \epsilon^t$ and $\epsilon^t\sim \N(0, \sigma^2)$, we know
\begin{equation}
g(\by|\bx) = (2\pi \sigma^2)^{-p/2} \exp(- \frac{\|\by - \bx + \alpha \nabla h(\bx)\|^2}{2\sigma^2}).
\end{equation}
Then the detailed balance condition is 
\begin{equation}\label{Eq:detailBalance}
p(\bx^{k+1}) g(\bx^{k}|\bx^{k+1}) = p(\bx^{k}) g(\bx^{k+1}|\bx^{k}).
\end{equation}
Since we will only consider $\bx^k$ and $\bx^{k+1}$ in the proof. For notation ease, we denote $\bx^{k+1}$ as $\bx^+$. 
In order to prove \ref{Eq:detailBalance}, we would like to show the ratio between LHS and RHS to be 1. To be more specific, we have (Recall that $\sigma^2 = \alpha$)
\begin{eqnarray}
\frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = & \exp(-h(\bx^{+}) - \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} + h(\bx^{k}) + \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha}).
\end{eqnarray}
Note that 
\[h(\bx^+) - h(\bx^k) = \frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + O(\|\bx^+ - \bx^k\|^2),
\]
and 
\begin{eqnarray}
 & \frac{\|\bx^{k} - \bx^+ + \alpha \nabla h(\bx^{+})\|^2}{2\alpha} - \frac{\|\bx^{+} - \bx^k + \alpha \nabla h(\bx^k)\|^2}{2\alpha} \\
 =& -\frac{\nabla h(\bx^k) + \nabla h(\bx^+)}{2} \cdot (\bx^+ - \bx^k) + \alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2.
\end{eqnarray}
Here, $\|\bx^+ - \bx^k\|$ will be roughly no bigger than $\alpha$ because $\bx^+ = \bx^k - \alpha \nabla h(\bx^k) + \epsilon^k$.  Thus it means $O(\|\bx^+ - \bx^k\|) = O(\alpha^2)$. What's more, $\alpha/2 \|\nabla h(\bx^+)\|^2 - \alpha/2 \|\nabla h(\bx^+)\|^2 = O(\alpha^2)$. 
Thus we know 
\begin{eqnarray}
\log \frac{p(\bx^{+}) g(\bx^{k}|\bx^{+})}{p(\bx^{k}) g(\bx^{+}|\bx^{k})} = &O(\alpha^2).
\end{eqnarray}
Therefore, we know the detailed balance equality holds approximately especially when $\alpha$ is small.

When $h$ is a quadratic function, the detailed balance holds exactly. 
\end{proof}

Some empirical demonstration. How accurate is this approximation?

@@

\begin{experiment}
Take the density function of $X$ to be $f(x) \propto \exp(- x^4)$, let $2\sigma^2 = \epsilon = 0.005$, $x_0 = 1$, burning-in period is $1000$, waiting period is $1000$, the empirical histogram v.s. the desired histogram is plotted in Figure ???
\end{experiment}
\begin{figure}
    \centering
    \includegraphics{figure/case1_step_0.005_iter_1e3.png}
    \caption{Experiment 1: the empirical distribution using SGD}
    \label{Fig:exp1}
\end{figure}

\section{Sampling from posterior distribution}
Consider an statistical model fitting problem. Given data $X_1,\ldots, X_n$, we are fitting a model $P(X_i|\theta)$. The maximum likelihood estimate will be $\hat{\theta} \in \argmin - \sum_{i=1}^n \log P(X_i|\theta)$. This can be solved by gradient descent on objective function $- \sum_{i=1}^n \log P(X_i|\theta)$. Now let's consider a Bayesian approach to estimate $\theta$ by maximizing its posterior probability given the data. The posterior distribution of $\theta$ is 
$P(\theta|(X_1,\ldots, X_n)) = \frac{\prod_{i=1}^n P(X_i|\theta) P(\theta)}{P(X_1,\ldots, X_n)} \propto \prod_{i=1}^n P(X_i|\theta) P(\theta)$. Then we could use SGD to draw sample from it. Note that we could approximate the noisy gradient step by subsampling. By central limit theorem, we know it still holds.

\begin{theorem}[Central limit theorem]

\end{theorem}

Suppose we would like to draw sample from a posterior distribution 
Some figures...
